\documentclass[12pt]{article}
\usepackage[width=16cm]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{aliases}
\usepackage{color}
\usepackage{url}

\usepackage{listings}
\usepackage{cancel}
\usepackage{textcomp}

\lstset{
   language=matlab,
   keywordstyle=\bfseries\ttfamily\color[rgb]{0,0,1},
   identifierstyle=\ttfamily,
   commentstyle=\color[rgb]{0.133,0.545,0.133},
   stringstyle=\ttfamily\color[rgb]{0.627,0.126,0.941},
   showstringspaces=false,
   basicstyle=\small,
   numberstyle=\footnotesize,
   numbers=none,
   stepnumber=1,
   numbersep=10pt,
   tabsize=2,
   breaklines=true,
   prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
   breakatwhitespace=false,
   aboveskip={0.1\baselineskip},
    columns=fixed,
    upquote=true,
    extendedchars=true,
% frame=single,
    backgroundcolor=\color[rgb]{0.9,0.9,0.9}
}

\title{One dimensional heat equation}
\author{Deep Ray, Ritesh Kumar, Praveen. C, Mythily Ramaswamy, J.-P. Raymond}
\date{}

\begin{document}



\maketitle
\begin{center}
IFCAM Summer School on Numerics and Control of PDE\\
22 July - 2 August 2013\\
IISc, Bangalore\\
\url{http://praveen.cfdlab.net/teaching/control2013}
\end{center}

%----------------------------------------------------------------------------------------

\section{The model}
The shifted 1D heat equation is given by
\[
z_t = \mu z_{xx} + \alpha z, \quad x \in (0,1) \times (0,T)
\]
with boundary conditions
\[
z(0,t) = 0, \quad z(1,t) = u(t)
\]
and initial condition
\[
z(x,0) = z_0(x) , \quad x \in (0,1)
\]
Here $\alpha \ge 0$ and $\mu > 0$. 
%----------------------------------------------------------------------------------------

\subsection{Weak formulation}
We assume $z_0 \in L^2(0,1)$. We wish to find $z \in L^2(0,T;H^1_{\{0\}}(0,1))$ such that
\[
 \dd{}{t}(z(t), \phi)_{L^2} = - \mu \int \limits^{1}_{0} \df{z}{x} \dd{\phi}{x} \, dx +  \alpha \int \limits^{1}_{0} z(x,t) \phi \, dx, \quad \forall \phi \in H^1_0(0,1)
\]
\[
 z(1,t) = u(t)
\]
\[
 (z(0),\phi)_{L^2} = (z_0 ,\phi)_{L^2}
\]
%----------------------------------------------------------------------------------------

\subsection{FEM approximation}
Consider a regular subdivision of $[0,1]$
\[
 0 = x_0 < x_1 < ... < x_N = 1, \quad h = \frac{1}{N}, \quad x_k = kh \quad \forall k \in \{0,1,...,N\}
\]
We work with the finite dimensional subspaces of $H^1_{0}(0,1)$
\begin{equation*}
\begin{aligned}
  & Z_h = \{ \phi \in C([0,1]) : \phi|_{[x_{i-1},x_i]} \in P_1, \quad \phi(0) = 0 \} \\
  & Z_{h,0} = \{ \phi \in C([0,1]) : \phi|_{[x_{i-1},x_i]} \in P_1, \quad \phi(0) = \phi(1) = 0 \}
\end{aligned}
\end{equation*}
We denote the standard $P_1$ bases for $Z_{h,0}$ and $Z_{h}$ by $\{\phi_1, \phi_2,...,\phi_{N-1}\}$ and $\{\phi_1, \phi_2,...,\phi_N\}$ respectively, with $\phi_j(x_i) = \delta_{ij}$. The test function are chosen from $Z_{h,0}$. The approximate weak formulation is given as
\[
 \text{Find } z = \sum \limits_{i=1}^{N-1} z_i \phi_i + u(t) \phi_N \in Z_h, \quad \text{ such that}
\]
\[
 \dd{}{t}(z(t), \phi_k)_{L^2} = - \mu \int \limits^{1}_{0} \df{z}{x} \dd{\phi_k}{x} \, dx +  \alpha \int \limits^{1}_{0} z(x,t) \phi_k \, dx, \quad \forall k \in \{1,..., N-1\}
\]
\[
 (z(0),\phi_k)_{L^2} = (z_0 ,\phi_k)_{L^2}, \quad \forall k \in \{1,..., N-1\}
\]
We obtain a a differential system 
\[
 \M \dd{\z}{t} = \A_\alpha \z + \B u , \quad \M \z(0) = ((z_0,\phi_i)_{L^2})_{1\le i \le {N-1}}
\]
where
\begin{itemize}
 \item $\z = (z_1,z_2,...,z_{N-1})^\top$
 \item $\M \in \re^{(N-1) \times (N-1)}, \quad M_{i,j} = \int \limits_0^1 \phi_i \phi_j \, dx$
 \item $\A \in \re^{(N-1) \times (N-1)}, \quad A_{i,j} = - \mu \int \limits_0^1 \dd{\phi_i}{x} \dd{\phi_j}{x} \, dx$
 \item $\A_\alpha = \A + \alpha \M$
 \item $\B =  \frac{\mu}{h} (0,0,...,0,1)^\top \in \re^{N-1}$
 \item $u \in \re$
\end{itemize}
Using exact integration for $\M$ and $\A$ gives
\[
 \M = h \left[\begin{array}{ccccc}
         \frac{2}{3} & \frac{1}{6} & 0 & \cdots & 0 \\
         \frac{1}{6} & \frac{2}{3} & \frac{1}{6}  & \vdots & 0 \\
         0 & \ddots  & \ddots & \ddots & 0 \\
         0 & \cdots & \frac{1}{6} & \frac{2}{3} & \frac{1}{6} \\
         0 & \cdots & 0 & \frac{1}{6} & \frac{2}{3}
        \end{array} \right] , \quad
 \A = \frac{\mu}{h} \left[\begin{array}{ccccc}
         \-2 & 1 & 0 & \cdots & 0 \\
         1 & -2 & 1  & \vdots & 0 \\
         0 & \ddots  & \ddots & \ddots & 0 \\
         0 & \cdots & 1 & -2 & 1 \\
         0 & \cdots & 0 & 1 & -2
        \end{array} \right]       
\]
We use the trapezoidal rule for $u'(t) \int_0^1 \phi_N \phi_{N-1}$ which will evaluate out as zero. 

\vspace{2mm}

\noindent
The matrices are computed in the matlab code {\tt matrix\_fem.m}
%----------------------------------------------------------------------------------------
\section{Time integration using BDF}
Assume for now that the feedback control is of the form 
\[
 u(t) = - \K \z(t), \quad \K \in \re^{1\times(N-1)}
\]
We shall use a Backward Differentiation Formula (BDF) for time integration. For the first time step we use BDF1 (backward Euler scheme)
\[
 \M \frac{\z^1 - \z^0}{\Delta t} = (\A_\alpha - \B \K)\z^1
\]
or
\[
 \left[ \frac{\M }{\Delta t} - (\A_\alpha - \B \K) \right] \z^1 =  \frac{\M}{\Delta t} \z^0 
\]
For the remaining time steps we use BDF2
\[
 \M \frac{\z^{n+1} - \frac{4}{3} \z^{n} + \frac{1}{3} \z^{n-1}}{\frac{2}{3}\Delta t} = (\A_\alpha - \B \K)\z^{n+1}
\]
or
\[
 \left[ \frac{\M }{\frac{2}{3}\Delta t} - (\A_\alpha - \B \K) \right] \z^{n+1} =  -\frac{\M}{\frac{2}{3}\Delta t}\left( -\frac{4}{3} \z^{n} + \frac{1}{3} \z^{n-1} \right)
\]

%--------------------------------------------------------------------------------
\section{Excercises}
\begin{enumerate}
\item The code in {\tt heat.m} solves the problem with homogeneous Dirichlet boundary conditions at $x=0$ and $x=1$ and initial condition
\[
z(x,0) = x^2 (1-x)^3
\]
Since $z(1,t) = u(t) = 0$ it is enough to solve
\[
\M \dd{\z}{t} = \A \z
\]
Run the code {\tt heat.m} with $\alpha =0$
\begin{lstlisting}
alpha = 0; 
mu = 1;
\end{lstlisting}
Are all the eigenvalues stable ? How do the solution and energy behave with time ?

\item Evaluate the decay rate of energy for initial condition $z(x,0) = \sin(\pi x)$
\begin{lstlisting}
P = polyfit(t, log(energy'), 1)
P(1) % Exact rate = -2*mu*pi^2
\end{lstlisting}
      
\item With initial condition $z(x,0) = \sin(\pi x)$ run {\tt heat.m} with
\begin{lstlisting}
alpha = 0.4 + pi^2 * mu
\end{lstlisting}
Are all the eigenvalues stable ? How do the solution and energy behave with time ?

\item For same $\alpha$ as above, compute 10 eigenvalues with largest real part using {\tt eigs} function. For each unstable eigenvalue $\lambda$, compute eigenvector $\V$ of $(\A_\alpha^\top,\M^\top)$
\[
 \A_\alpha^\top \V = \lambda \M^\top \V
\]
and check if 
\[
 \B^\top \V \ne 0
\]
If the above is true for all unstable eigenvalues, then the system is stabilizable. Is the heat equation system stabilizable ?

\item Copy {\tt heat.m} as {\tt heat1.m} and modify {\tt heat1.m} to implement the following non-homogeneous Dirichlet boundary condition
\[
z(1,t) = u(t) = 1
\]
and use the same initial condition as before. This means that the code must now solve $\M \dd{\z}{t}=\A \z + \B$. Use $\mu=1$ and $\alpha=0$. The solution should converge to
\[
\lim_{t \to \infty} z(x,t) = x
\]

\end{enumerate}

%--------------------------------------------------------------------------------
\section{Minimal norm feedback control}
The minimal norm control is given by
\[
u(t) = - \K z (t)
\]
The feedback matrix $\K$ is given by
\[
\K = \R^{-1} \B^\top \bdP \M
\]
where $\bdP$ is solution of algebraic Riccati equation (ARE)
\[
\A^\top_\alpha \bdP \M + \M^\top \bdP \A_\alpha - \M^\top \bdP \B \R^{-1} \B^\top \bdP \M  = 0
\]
This can be solved using the {\tt care} function
\begin{lstlisting}
[P,L,K] = care(A,B,R,Q,S,M);
\end{lstlisting}
which solves the more general ARE
\[
\A^\top \bdP \M + \M^\top \bdP \A - (\M^\top \bdP \B \ + \bdS^\top)  \R^{-1} (\B^\top \bdP \M +\bdS) + \Q = 0
\]
where $\bL$ gives the eigenvalues of $(\A- \B \K,\M)$. Since $\bdS=0$ for our problem, we can simply replace it with {\tt []} while calling the {\tt care} function. Furthermore, for the minimal norm feedback, we take $\Q=0$, which can be set in {\tt matrix\_fem.m}
%--------------------------------------------------------------------------------
\subsection{Excercises}

\begin{enumerate}
\item Set $\Q=0$ in {\tt matrix\_fem.m} which gives minimal norm control and use
\begin{lstlisting}
alpha = 0.4 + pi^2 * mu
\end{lstlisting}
Run the code {\tt heat\_lqr.m}. Use initial condition $z(x,0) = \sin(\pi x)$. Observe the modification in the eigenvalues; zoom the eigenvalue plot near the origin and locate the unstable eigenvalue and modified eigenvalue; note that the unstable eigenvalue is reflected about the imaginary axis. What about the remaining stable eigenvalues? Is the solution of the heat equation stable ?

\item Save the feedback matrix evaluated
\begin{lstlisting}
save('feedback.mat','K')
\end{lstlisting}
Copy {\tt heat.m} as {\tt heat2.m}. Load $\K$ into {\tt heat2.m}, and for {\tt alpha = 0} (no shift) introduce control using the feedback matrix
\begin{lstlisting}
load('feedback');
A = A-B*sparse(K);
\end{lstlisting}
We now solve $\M \dd{\z}{t}=(\A - \B\K)\z$. Find the decay rate in this case using {\tt polyfit} function. We expect it to be atleast $2(0.4 + \mu \pi^2)$.

\end{enumerate}

%--------------------------------------------------------------------------------

\section{Feedback control using LQR approach}
Performance measure
\begin{eqnarray*}
J &=& \frac{1}{2}\int_0^\infty \int_0^1 z^2(x,t) \ud x \ud t + \frac{1}{2} \int_0^\infty u^\top R u \ud t \\
&=& \frac{1}{2}\int_0^\infty \z^\top \Q \z \ud t + \frac{1}{2} \int_0^\infty u^\top \R u \ud   
t , \qquad \Q = \M = \C^\top \C, \qquad \C = \M^\frac{1}{2}
\end{eqnarray*}
The feedback law
\[
u = -\K \z
\]
which minimizes $J$ is given by
\[
\K = \R^{-1} \B^\top \bdP \M
\]
where $\bdP$ is solution of algebraic Riccati equation (ARE)
\[
\A^\top_\alpha \bdP \M + \M^\top \bdP \A_\alpha - \M^\top \bdP \B \R^{-1} \B^\top \bdP \M + \Q = 0
\]
%The solution exists if $(A,B)$ is controllable and $(A,C)$ is observable.
%--------------------------------------------------------------------------------

\subsection{Excercises}

\begin{enumerate}

\item In {\tt matrix\_fem.m}, put $\Q = \M$. Run {\tt heat\_lqr.m}. How do the eigenvalues change in LQR as compared to minimal norm control ? Is the solution stable with LQR feedback ?

\item Plot the intial condition and the numerical solution after one time step. How does it differ from the previous Bernoulli (minimal norm) control case ?

\item How does the solution and control vary as {\tt mu} is decreased ?

\item In {\tt matrix\_fem.m}, vary the value of $\R$ in (0.01, 50) (choose a couple of values in the range). How does $\K$ change with $\R$ ?

\item As before, save the feedback matrix and load it into {\tt heat.m}. Is the decay rate increased with LQR feedback as compared to case of minimal norm control ?
\end{enumerate}

%--------------------------------------------------------------------------------

\section{Control based on stabilizing only the unstable components}
Let the us denote by $\lambda_i$ and $\V_i$ the eigenvalues and normalized eigenvectors such that
\[
 \A \V_i = \lambda_i \M \V_i
\]
We assume that $\alpha$ is chosen such that
\[
 \lambda_{N-1} < .... <\lambda_{N_\alpha +1} < -\alpha < \lambda_{N_\alpha} < ... < \lambda_1
\]
Thus in the shifted system for the heat equation, there will be $N_\alpha$ unstable eigenvalues.
\[
 \lambda_{N-1} + \alpha < .... <\lambda_{N_\alpha +1} + \alpha < 0 < \lambda_{N_\alpha} + \alpha < ... < \lambda_1 + \alpha
\]
The eigenvectors $\V_i$ form a basis for $\re^{N-1}$ and have the property
\[
 \V_i^\top \M \V_j = \delta_{i,j}, \quad \V_i^\top \A \V_j = \delta_{i,j} \lambda_i
\]
Let $\Sigma \in \re^{(N-1) \times (N-1)}$ with $\V_i$ forming the columns, and consider the variable change
\[
 \z = \Sigma \zz
\]
Thus the shifted system can be written as
\[
 \dd{\zz}{t} = \Lambda_\alpha \zz + \BB u
\]
where
\[
 \Lambda_\alpha = \begin{bmatrix}
                   \alpha + \lambda_1 & 0 & \cdots & 0\\
                   0 & \alpha + \lambda_2 & \cdots & 0\\
                   \vdots & \vdots & \ddots & &\\
                   0 & 0 & 0 & \alpha + \lambda_{N-1}
                  \end{bmatrix}, \qquad
 \left[\BB\right]_i = V^\top_i \B, \qquad  1 \le i \le N-1                                   
\]
Next we project the system onto the unstable subspace using the projection operator $\PIu \in \LL(\re^{N-1},\re^{N_\alpha})$
\[
 \dd{\zzu}{t} = \Lambda_{\alpha,u} \zzu + \BBu u
\]
where
\[
 \Lambda_{\alpha,u} = \begin{bmatrix}
                   \alpha + \lambda_1 & 0 & \cdots & 0\\
                   0 & \alpha + \lambda_2 & \cdots & 0\\
                   \vdots & \vdots & \ddots & &\\
                   0 & 0 & 0 & \alpha + \lambda_{N_\alpha}
                  \end{bmatrix}, \qquad
 \left[\BBu \right]_i = V^\top_i \B, \qquad 1 \le i \le N_\alpha                                   
\]
and $\zzu = \PIu \zz$ are the first $N_\alpha$ components. We find the feedback matrix for this reduced system, by solving the Bernoulli equation
\begin{equation}
 \PPu \Lambda_{\alpha,u} + \Lambda^\top_{\alpha,u} \PPu - \PPu \BB_u \BB_u^\top \PPu = 0
\label{eq:ricu}
\end{equation}
\[
 \Lambda_{\alpha,u} - \BBu \BBu^\top \PPu \text{ is stable}
\]
The corresponding matrix $\bdP \in \LL(\re^{N-1})$ such that $(\M,\A_\alpha - \B\B^\top\bdP\M)$ is stable is given by
\[
 \bdP = \Sigma_u \PPu \Sigma^\top_u
\]
where $\Sigma_u$ is the matrix of eigenvectors corresponding to the unstable eigenvalues.
\[
\Sigma_u = [ V_1, V_2, \ldots, V_{N_\alpha}]
\]
This is implemented in {\tt heat\_lqru.m}
%----------------------------------------------------------------------------------------

\subsection{Excercises}

\begin{enumerate}

\item In {\tt matrix\_fem.m}, put $\Q = 0$. Run {\tt heat\_lqru.m}. How do the eigenvalues change this time  compared to minimal norm control ? Is the solution stabilized ?

\item Plot the intial condition and the numerical solution after one time step. How does it differ from the previous two kinds of control ?

\item For the same value of {\tt alpha = 0.4 + mu * pi * pi}, which of the three types of controls discussed so far performs the best, i.e., which one gives th fastest stabilization ?

\item Lecture 5 has the exact solutions for this problem, with {\tt alpha = 10}. Take 
\begin{lstlisting}
 z0 = sin(pi*x) 
\end{lstlisting}
in the equation and find the expressions for exact control and solution. Plot and compare 
\begin{itemize}
 \item evolution of the numerical control and the exact control 
 \item numerical and exact solutions at the final time
\end{itemize}

\item Save the feedback matrix and load it into {\tt heat.m}. Is the decay rate increased with feedback ? How does it compare with the case where we solve the full Riccati equation as in previous section ?

\item This approach is efficient since we only have to solve a small dimensional Riccati equation. Try to solve the full Riccati equation with a large number of elements, say {\tt ni = 500} and/or {\tt ni = 1000} and find the time taken by matlab. Compare this with the time taken to solve the Riccati equation~(\ref{eq:ricu}) whose dimension is small and independent of the number of elements.
\end{enumerate}

\pagebreak
%-----------------------------------------------------------------------------------

\section{System with noise and partial information}
Consider the system with noise in the model and initial condition
\[
z_t = \mu z_{xx} + \alpha z + w, \qquad z(x,0) = z_0(x) + \eta
\]
where $w$ and $\eta$ are error/noise terms. The boundary conditions are as before
\[
z(0,t) = 0, \quad z(1,t) = u(t)
\]
We may not have access to the full state but only some partial information which is also corrupted by noise.
\[
y = Hz  + v
\]
where $H$ is a suitable measure. We shall consider the case where $H$ is given by
\[
 Hz(t) = \mu z_x(0,t)
\]
We have to approximate $z_x(0,t)$ by a finite difference formula. We can use a first order or second order finite differences
\[
\begin{aligned}
\mbox{first order} \qquad & \mu z_x(0,t_n) \approx \mu \frac{z_1^n - z_0^n}{h} = \mu \frac{z_1^n}{h} \\
\mbox{second order} \qquad & \mu z_x(0,t_n) \approx \mu \frac{- 3 z_0^n + 4 z_1^n - z_2^n}{2h} = \mu \frac{4 z_1^n - z_2^n}{2h}
\end{aligned}
\]
In the FEM setup, we get the system
\[
 \M \dd{\z}{t} = \A_\alpha \z + \B u + \w , \quad \M \z(0) = ((z_0,\phi_i)_{L^2})_{1\le i \le {N-1}}
\]
\[
 \y = \bH \z + \bdv
\]
where 
\[
 \bH = \frac{\mu}{h}(2,0.5,0,...,0)^\top \in \re^{N-1}
\]


%--------------------------------------------------------------------------------

\subsection{Estimation problem}
Linear estimator
\[
\dd{z_e}{t} = A_\alpha z_e + B u + L(y - H z_e)
\]
We determine the filtering gain $L$ by minimizing
\[
J = \frac{1}{2} \int_0^\infty (y - H z_e)^\top R^{-1}_v (y - H z_e) \ud t +    \frac{1}{2} \int_0^\infty w^\top R_w^{-1} w \ud t
\]
The solution is given by
\[
L = P_e H^\top R^{-1}_v
\]
where $P_e$ is solution of
\[
A_\alpha P_e + P_e A_\alpha^\top - P_e H^\top R_v^{-1}  H P_e + R_w = 0
\]
The operators $R_w$ and $R_v$ are chosen according to the apriori knowledge we have on the model noise and the measurement noise. In the FEM setup, we have
\[
\M \dd{\z_e}{t} = \A_\alpha \z_e + \B u + \bL(\y - \bH \z_e)
\]
\[
\bL = \M \bdP_e \bH^\top \R^{-1}_\bdv
\]
where $\bdP_e$ is solution of
\[
\A_\alpha \bdP_e \M + \M \bdP_e \A_\alpha^\top - \M \bdP_e \bH^\top \R_\bdv^{-1}  \bH \bdP_e \M + \R_\w = 0
\]
\[
 (\M,\A_\alpha - \bL \bH) \text{ is stable} 
\]

%--------------------------------------------------------------------------------

\subsection{Coupled linear system}
The feedback is based on estimated solution $u = -\K \z_e$
\begin{eqnarray*}
\M \dd{\z}{t} &=& \A_\alpha \z - \B \K \z_e + \w \\
\M \dd{\z_e}{t} &=& \bL \bH \z + (\A_\alpha - \B \K - \bL \bH) \z_e + \bL \bdv
\end{eqnarray*}
or in matrix form
\[\begin{bmatrix}
   \M & 0\\
   0 & \M \\
  \end{bmatrix}
\dd{}{t} \begin{bmatrix}
\z \\
\z_e \end{bmatrix} = \begin{bmatrix}
\A_\alpha & -\B \K \\
\bL \bH & \A_\alpha - \B \K - \bL \bH \end{bmatrix} \begin{bmatrix}
\z \\ \z_e \end{bmatrix} + \begin{bmatrix}
\I & 0 \\
0 & \bL \end{bmatrix} \begin{bmatrix}
\w \\ \bdv \end{bmatrix}
\]
The initial condition is given by
\[
\z(0) = \z_0 + \eta, \qquad \z_e(0) = \z_0
\]
This is implemented in program {\tt heat\_est.m}
%----------------------------------------------------------------------------------------

\subsection{Excercises}

\begin{enumerate}

\item Study the program {\tt heat\_est.m} and run it. We are taking initial condition
\[
\z(0) = \sin(\pi x) + 0.1 \sin(2 \pi x), \qquad \z_e(0) = \sin(\pi x)
\]

\item Look at eigenvalues of $(A,M)$, $(A-BK,M)$ and $(A-LH,M)$. Are they stable?

\item By default, the program takes $\Q=0$. Try the LQR approach using $\Q = \M$.

\end{enumerate}

%--------------------------------------------------------------------------------
\subsection{Stabilizing the unstable component}

We use the variable change to the eigenvector basis
\[
 \z = \Sigma \zz
\]
The first $N_\alpha$ eigenvalues are unstable; the projection to the unstable space is $\zzu = \PIu \zz$. As done before, the feedback gain matrix $\K$ can be evaluated by considering the projected system
\[
 \dd{\zzu}{t} = \Lambda_{\alpha,u} \zzu + \BBu u
\]
Define the operator $\HH$ and $\HHu$ by
\[
 \HH \zz = \bH \z, \qquad \HHu = \HH \PIu
\]
The filtering gain $\PP_e$ for the reduced projected system is obtained as the solution of the ARE
\[
 \PP_e \Lambda_{\alpha,u}^\top + \Lambda_{\alpha,u} \PP_e - \PP_e \HHu^\top \R_\bdv^{-1} \HHu \PP_e + \QQ_\w = 0
\]
\[
 \Lambda_{\alpha,u} - \PP_e \HHu^\top \R_\bdv^{-1} \HHu \text{ is stable} 
\]
where $\QQ_\w = \Sigma_u^\top \R_\w \Sigma_u$. The corresponding $\bdP_e$ such that $(\M,\A_\alpha - \M \bdP_e \bH^\top \R_\bdv^{-1} \bH)$ is stable is given by
\[
 \bdP_e = \Sigma_u \PP_e \Sigma^\top_u
\]

%--------------------------------------------------------------------------------
\subsection{Excercises}

\begin{enumerate}

\item Write a code {\tt heat\_estu.m} for the above via the following steps.

\begin{itemize}
\item Copy the contents of {\tt heat\_est.m} to a new file called {\tt heat\_estu.m}

\item Evaluate the feedback gain $\K$ as done in {\tt heat\_lqru.m}

\item Evaluate the projected matrix $\HHu$ 
\begin{lstlisting}
Hu = full(H*V); 
\end{lstlisting}
where {\tt V} is the matrix of unstable eigenvectors (you would have already evaluated this while obtaining $\K$).

\item Evaluate $\QQ_\w$ 
\begin{lstlisting}
Rwu = full(V'*Rw*V); 
\end{lstlisting}

\item Solve the ARE for the reduced system to obtain $\PP_e$
\begin{lstlisting}
Peu = care(D',Hu',Rwu,Rv);
\end{lstlisting}
where {\tt D} is the diagonalized matrix of the reduced system ( also evaluated while finding $\K$ )

\item Find $\bdP_e$ for the original system
\begin{lstlisting}
Pe = sparse(V*Peu*V'); 
\end{lstlisting}

\item Finally evaluate $\bL$ by
\begin{lstlisting}
L = M*Pe*H'/Rv; 
\end{lstlisting}

\item Run {\tt heat\_estu.m} with {\tt alpha = 10}. How do the solution and energy behave~?

\end{itemize}

\end{enumerate}

%--------------------------------------------------------------------------------

\section{List of Programs}

\begin{enumerate}

\item {\tt matrix\_fem.m}: Computes FEM matrices

\item {\tt feedback\_matrix.m}: Computes feedback matrix

\item {\tt heat.m}: Solves the heat equation for a given $\alpha$, without feedback

\item {\tt heat\_lqr.m}: Solves the heat equation for a given $\alpha$, with feedback

\item {\tt heat\_lqr.m}: Solves the heat equation for a given $\alpha$, stabilizing only the unstable components

\item {\tt heat\_est.m}: Solves the coupled estimation and control problem for the heat equation for a given $\alpha$

\item {\tt heat\_estu.m}: Solves the coupled estimation and control problem for the heat equation for a given $\alpha$, stablilizing only the unstable components.

\end{enumerate}

\end{document}  
